Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes

### 2020-05-28 Ben_Manuscripts/hdphmm/Draft.tex
Comments.

  We can better understand the mechanisms controlling these dynamical modes by grouping 
  the states identified by the HDP-AR-HMM into clusters based on multiple metrics
  aimed at distinguishing solute behavior based on their fluctuations, dwell times
  in each state and position within the inhomogeneous membrane structure. We analyze
  prevalent clusters in order to relate their parameters to physical interactions between 
  solutes and the membrane.

@@

  One issue with the original formulation of the HDP-HMM is that it does 
  not distinguish self-transitions and transitions between different states.
  Practically, this can result in rapid switching between states and inferred
  state sequence representations that do a poor job of describing the 
  observations.
  %BJC: added by Chris  
  Rapid state switching in the original HDP-HMM not only reduces the predictive
  ability of the model, but it is also physically unrealistic and can complicate
  model interpretation. 
  Fox et al. addressed this issue by adding a `sticky' 
  parameter.~\cite{fox_sticky_2007} The sticky HDP-HMM is useful in situations
  where the user has prior knowledge that state persistence is likely.
  For example, in the case of speaker diarization, it is unlikely that a 
  conversation between two people would rapidly switch between speakers.~\cite{fox_sticky_2011}
  This is also important for our work where we expect state persistence due
  to solute entrapment.

@@

  % BJC: fair number of changes to this paragraph based on Chris' suggestions.
  % BJC: he really wants use to emphasize our contribution with this part
  To our knowledge, applying the HDP-AR-HMM in order to gain kinetic and
  mechanistic insight from MD simulation trajectories has not been done
  before. Typical applications of the HDP-AR-HMM center around determining the 
  number of states, the state sequence and analysis of the emission 
  distributions of a time series. The addition of temporal correlation to 
  the inference procedure has primarily been used to improve state sequence 
  identification.~\cite{calderon_inferring_2015} In our approach, we use the
  HDP-AR-HMM not only to segment a collection of MD time series into distinct 
  dynamical modes, but to uncover unknown transport mechanisms and create 
  improved predictive models. We cluster the parameters of individual states
  in order to reduce the total state space down to an interpretable size
  This allows us to learn the implied dominant physical mechanisms and describe
  solute kinetics based on interpretation of distinct parameters that dominate
  the ensemble of MD trajectories. We can use the model parameters determined 
  either before or after clustering to generate new time series trajectories 
  which enable projections of the MD-generated time series to arbitrary length
  time scales.

@@

  It is usually beneficial to employ intelligent state sequence initialization procedures
  in nonparametric Bayesian approaches.~\cite{fox_bayesian_2010}
  Most importantly in this work, the states identified by the HDP-AR-HMM are heavily 
  influenced by the Gaussian prior placed on $\mathbf{c}$ in Equation~\ref{eqn:var}. In 
  Section~\ref{S-section:fitting_tips} of the Supporting Information we outline our 
  algorithmic method for choosing reliable prior parameters in addition to other 
  practical tips for initialization.

  %BJC: The following paragraph used to be the first paragraph after italics above. 
  % I moved it down here as suggested by Chris below. What do you think in terms of flow?  
  We inferred the state sequence and parameters for each of the 24 solute 
  trajectories independently. Although the HDP-AR-HMM is capable of identifying 
  an infinite number of states, a Dirichlet process tends to exhibit a 
  ``rich get richer" effect, favoring a fewer number of states.~\cite{dreyer_discovering_2011}
  By applying the algorithm to each trajectory independently, we reduce the 
  possibility of lumping together multiple similar states which we would prefer
  to stay separated before clustering. 
  %BJC: added following sentence to aid reader in response to Chris' question below 
  In Section~\ref{method:realizations}, we 
  explain how we use these 24 independent models in order to generate a single
  MSD prediction.
  % BJC: abridged Chris comment with respect to this paragraph here:
  % CPC: the question i have at this point as a reader in this area, given 
  % how this is setup both as predictive and clustering, is "how do they plan to
  % generate data and predict with this model if they have 24 different transition
  % mat estimates"
  % CPC: returning to comment after reading.  section 2.4 explains what you do 
  % (though i'm still fuzzy on how you estimate transition matrix probs in method 2). 
  % (BJC: Note I tried to clarify this process for method 2 below)
  % CPC_continued: in any case, you might want to avoid obvious critiques and head 
  % scratching from expert reviewers. i'd suggest moving this paragraph to end of subsection and
  % expand on above considerations after you've given more cut and dry details below and reference 2.4 stating     
  %that you say how you simulate from a model estimated one trajectory at a time via HDP-AR-HMM inference

@@

  We clustered states with like parameter sets in order to reduce the state space to
  a more easily interpretable size 
  %BJC: added following in response to Chris' below comment. I think I'm interpretting what he's saying right
  % CPC: perhaps keep hammering home not only reducing, but also reemphasize why absolute
  % value of some parameters really not of interest 
  and because the parameters of a single states does not always give meaningful physical insight. 
  For each solute studied, we identified approximately 300-500 independent states, 
  each with separate VAR(1) parameters. Many of these states exhibit very similar
  dynamical behavior except their mean levels in each dimension are different.

@@

  We obtained estimates of clustered $A$ and $\Sigma$ parameters as well as a
  single $T$, describing transitions between all possible clustered states exhibited by
  the ensemble of trajectories, using the inference component of the HDP-AR-HMM
  algorithm. We applied the HDP-AR-HMM simultaneously on all 24 zeroed solute
  trajectories with state sequences fixed according to the remapped clustered 
  state assignments. This is only possible because the state sequence is fixed. 
  In cases where one does not know the state sequence and would like to 
  incorporate data from multiple trajectories, one can apply the beta process (BP)
  AR-HMM.~\cite{fox_sharing_2009} However the data requirements may be quite high
  for a good parameterization, which is why we elected to follow the clustering 
  approach described thus far.
  %BJC: I actually found 3 different ways do this. All giving the same results.
  % Method 1: described above - this is the most efficient way
  % Method 2: Link all 24 trajectories by adding a coffin state (basically a 
  % state that is a vector of zeros) between each traj
  % Method 3: Parameterize each trajectory independently, average together. T is
  % estimated by making a count matrix based on inputs from each trajectory

@@

  For both approaches, we drew state sequences with transition probabilities given 
  % BJC: Chris wasn't clear where T came from for method 2. I tried to make it clearer in 
  % last paragraph of previous section. Let me know if it's not
  by $T$ of the HDP-AR-HMM. While in a given state, we simulated motion according to the 
  VAR(1) parameterization of that state. After each state transition, we set the 
  unconditional mean of the new state to the particle's position immediately before
  the state transition occurred.

@@

  Solute flux, as written, can be expressed in terms of the product of solute 
  diffusivity, $D_i$, within the membrane and a sorption coefficient, $K_i$, 
  that quantifies the ability of the solute to move between the bulk solution
  and and the membrane:
  \begin{equation}
    J_i = \frac{D_i K_i}{l}\Delta C
  \end{equation}
  where $l$ is the membrane thickness. For our purposes, $K_i$ and $l$ are 
  irrelevant since we do not study pore entry effects or flux across finite
  length pores. Therefore, in order to calculate selectivity, with respect
  to quantities we can study with our MD simulations, it is only necessary that
  we compare the ratio of diffusion constants:
  \begin{equation}
    S_{ij} = \frac{D_i}{D_j}
    \label{eqn:selectivity_diffusivity}
  \end{equation}
  It can be shown on time scales where motion is uncorrelated, that the diffusion
  constant is proportional to the slope of the linear region of the MSD 
  curve:~\cite{einstein_investigations_1956}
  \begin{equation}
    \overline{z^2(\tau)} = 2 n D t
    \label{eqn:msd_D}
  \end{equation}
  where $n$ is the number of spatial dimensions (1 in our case). In general, the 
  MSD of the realizations of our models enters a linear regime near a 5 ns time lag
  (see Figure~\ref{S-fig:MET_nclusters} of the Supporting Information for example).
  Since MSD sampling decreases with increasing time lag, longer trajectory 
  realizations help to extend the linear region.
  %BJC: need to confirm these details
  For all solutes, we generated 200 ensembles of 24 5 $\mu s$ trajectories and 
  estimated $D$ from the slope fit between 10 ns and 1000 ns. We used weighted 
  least squares to fit the slope with weights determined by the number of time 
  lags used to compute the average ($T - \tau$ in equation~\ref{eqn:tamsd}). 
  We bootstrapped error bars for the final reported selectivities. For each 
  bootstrap trial, we sampled the 200 trajectory ensembles with replacement, 
  averaging the mean MSD curves calculated from each ensemble and then calculating its 
  slope to create a distribution of $D$ values. We report the mean and standard
  error of the distribution of $D$.

@@

  %BJC6: added. Should I add a figure?
  A potential shortcoming of the HDP-AR-HMM applied to MD simulations is its 
  inability to reproduce the short time scale MSD curves. The MSDs predicted
  by the model tend to enter a linear regime much more quickly than the MD MSDs
  (see Figure~\ref{S-fig:unclustered_msd_shortlag} of the Supporting Information
  for example). This is not a major concern for predicting macroscopic properties
  since we are primarily interested in reproducing the MSDs on much longer 
  timescales.
  %BJC6: could hypothesize why using previous work as support

@@

  The data suggests that the particular LLC membrane studied in this work is
  capable of selectively separating urea and acetic acid from ethylene glycol
  and methanol. Ethylene glycol and methanol have comparable diffusivities
  while urea and acetic acid have much lower diffusivities.

@@

  The parameters of the five dominant states provide clues about the associated solute
  behavior. Characteristic dynamics of the following states appear in 
  Figure~\ref{fig:clustered_traj_MET}. A more in-depth analysis of the state parameters
  is given in Section~\ref{S-section:clustered_parmeters} of the Supporting Information.
  \begin{enumerate}[label={State \theenumi :}, leftmargin=3.5\parindent]
     \item This state has intermediate dwell times with high past-hop dependence (high
     entries of $A$). The cluster's radial mean is 1.5 nm which is about halfway between
     the pore center and tail ends. Evidence from Figure~\ref{fig:clustered_traj_MET} 
     suggests that this state occurs essentially evenly across possible radial
     distances from the pore center.
     \item This state's radial mean of 2.4 nm implies that it occurs exclusively in the
     tail region based. Its fluctuations are small and there is little past hop dependence.
     \item This is likely a hopping state that strongly contributes to the solute's MSD. 
     It is short lived and has large fluctuations which often results in very large hops.
     Like State 1, we see this type of behavior near the pore center and near the ends of the tails.
     \item This is a trapping state, having the longest expected dwell time among the dominant
     states, with a strong dependence on its previous fluctuation. Evidence from 
     Figure~\ref{fig:clustered_traj_MET} and a radial mean of 1.8 suggest that it tends to
     occur at intermediate distances from the pore center with perhaps a higher tendency
     to occur deeper within the tails.
     \item This is an intermediate-length trapping state that occurs 0.9 nm from the pore 
     center on average. It's fluctuations are tight and relatively independent of its previous 
     fluctuation.
  \end{enumerate}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{hbonds_assoc_summary.pdf}
  \caption{}\label{fig:hbonds_assoc_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{rdf_summary.pdf}
  \caption{}\label{fig:rdf_summary}
  \end{subfigure}
  %BJC: Is it worth putting a bars with the solute MSDs on this plot
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{dwell_time_summary.pdf}
  \caption{}\label{fig:dwell_time_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{density_comparison.pdf}
  \caption{}\label{fig:density_comparison}
  \end{subfigure}
  \caption{(a) Solutes hydrogen bond and associate with sodium ions
  to different degrees. Sometimes, they participate in both interactions at
  once. (b) We histogrammed the radial means of the unclustered states weighted
  by the length of time spent in that state. Most solutes spend the majority of
  their time close to the pore center. Only methanol has a significant density
  far from the pore center. (c) The largest expected value (see 
  Equation~\ref{eqn:dwell_times}) of dwell times among the dominant clustered 
  states for each solute are inversely related to the solute's MSD. 
  % BJC: Copying Chris' comments below:
  % CPC: Discussion is confusing. guessing error bars are from measured simulation 
  % dwell times in dominant state.  
  % BJC: will add error bar explanation once I figure out exactly how to modify the figure
  % CPC: what if trajectory has more than one dominant state? (seems from fig like 
  % you take one with largest dwell time from your model. 
  % BJC: All solutes have at  least 4 dominant states by my definition. maybe it would 
  % be better to sample dwell times from all of the dominant states. Just weight 
  % their contributions. Not sure if weighting should be with respect to the fraction
  % the total simulation time spent in each state, or the number of times the state
  % was visited.
  % CPC: might want to better connect MSDs of model in fig.3 to fig 9 c and 
  % expand/clarify discussion on connection of fig 9c to other panels just lots
  % of ambiguity here, but results seem promising.
  % BJC: is it worth plotting bars with the MSD after 1000 ns as part of (c)? 
  %MRS5: not clear what ``largest expected'' means here.
  %BJC6: modified to 'expected value' and referenced equation   
  (d) The expected dwell time of solutes increases 
  in states with high local number density of surrounding heavy atoms.
  }\label{fig:summaries}
  \end{figure}

  % BJC: Chris was confused by this paragraph
  Solutes with dominant states characterized by long dwell times tend to 
  have lower MSDs. We define the dominant states exhibited by each solute
  as any state visited by at least half (12) of the solute trajectories.
  By our definition and based on our choice to study 10 clusters for each 
  solute, methanol, ethylene glycol, urea and acetic acid are characterized
  %BJC: added this supporting figure 
  by 5, 6, 5 and 4 dominant states respectively (see Figure~\ref{fig:state_prevalence}
  of the Supporting Information). In Figure~\ref{fig:dwell_time_summary}, we
  plot the expected dwell time of each solute in the dominant state which 
  has the the highest self-transition probability. The MSDs shown in 
  Figure~\ref{fig:unclustered_msds} are inversely related to the maximum 
  expected dwell times. Both acetic acid and urea participate in many hydrogen
  bonds and sodium ion association interactions and the states visited by both 
  solutes have similar distributions of radial means, however acetic acid is 
  subject to much longer dwell times, and hence a lower MSD. This may be 
  because acetic acid is a stronger hydrogen bond donor than urea.

@@

  When clustering, our goal was to create the simplest model possible that gave
  clear mechanistic insight. This is important for systems where transport mechanisms
  are not easily hypothesized. However, once one gains some mechanistic understanding,
  as we have in this work, the clustering procedure can be modified in order to 
  study specific interactions or radially dependent behavior.

>>>>

  We can better understand the mechanisms controlling these dynamical modes by grouping 
  the states identified by the HDP-AR-HMM into clusters based on multiple metrics
  aimed at distinguishing solute behavior based on their fluctuations, dwell times
  in each state and position within the inhomogeneous membrane structure. We analyze
  prevalent clusters in order to relate their 
  %MRS: suggestion to add:
  dynamical
  parameters to physical interactions between 
  solutes and the membrane.

@@

  One issue with the original formulation of the HDP-HMM is that it does 
  not distinguish self-transitions and transitions between different states.
  Practically, this can result in rapid switching between states and inferred
  state sequence representations that do a poor job of describing the 
  observations.
  %BJC: added by Chris  
  Rapid state switching in the original HDP-HMM not only reduces the predictive
  ability of the model, but 
  %MRS: tweaked; sometimes it might be physical.
  % it
  in many cases
  is also physically unrealistic and can complicate
  model interpretation. 
  Fox et al. addressed this issue by adding a `sticky' 
  parameter.~\cite{fox_sticky_2007} The sticky HDP-HMM is useful in situations
  where the user has prior knowledge that state persistence is likely.
  For example, in the case of speaker diarization, it is unlikely that a 
  conversation between two people would rapidly switch between speakers.~\cite{fox_sticky_2011}
  This is also important for our work where we expect state persistence due
  to solute entrapment.

@@

  % BJC: fair number of changes to this paragraph based on Chris' suggestions.
  % BJC: he really wants us to emphasize our contribution with this part
  To our knowledge, applying the HDP-AR-HMM in order to gain kinetic and
  mechanistic insight from MD simulation trajectories has not been done
  before. Typical applications of the HDP-AR-HMM center around determining the 
  number of states, the state sequence and analysis of the emission 
  distributions of a time series. The addition of temporal correlation to 
  the inference procedure has primarily been used to improve state sequence 
  identification.~\cite{calderon_inferring_2015} In our approach, we use the
  HDP-AR-HMM not only to segment a collection of MD time series into distinct 
  dynamical modes, but to uncover unknown transport mechanisms and create 
  improved predictive models. We cluster the parameters of individual states
  in order to reduce the total state space down to an interpretable size
  This allows us to learn the implied dominant physical mechanisms and describe
  solute kinetics based on interpretation of distinct parameters that dominate
  the ensemble of MD trajectories. We can use the model parameters determined 
  either before or after clustering to generate new time series trajectories 
  which enable projections of the MD-generated time series to arbitrary length
  time scales.

@@

  It is usually beneficial to employ 
  %MRS: tweak
  %intelligent 
  well-informed
  state sequence initialization procedures
  in nonparametric Bayesian approaches.~\cite{fox_bayesian_2010}
  Most importantly in this work, the states identified by the HDP-AR-HMM are heavily 
  influenced by the Gaussian prior placed on $\mathbf{c}$ in Equation~\ref{eqn:var}. In 
  Section~\ref{S-section:fitting_tips} of the Supporting Information we outline our 
  algorithmic method for choosing reliable prior parameters in addition to other 
  practical tips for initialization.

  %BJC: The following paragraph used to be the first paragraph after italics above. 
  % I moved it down here as suggested by Chris below. What do you think in terms of flow?  
  We inferred the state sequence and parameters for each of the 24 solute 
  trajectories independently. Although the HDP-AR-HMM is capable of identifying 
  an infinite number of states, a Dirichlet process 
  %MRS: tweak
  %tends to exhibit a 
  exhibits a 
  ``rich get richer" effect, favoring a fewer number of states.~\cite{dreyer_discovering_2011}
  %MRS: this next sentence is a bit vauge.  Why would we prefer thtem to stay separated?
  By applying the algorithm to each trajectory independently, we reduce the 
  possibility of lumping together multiple similar states which we would prefer
  to stay separated before clustering. 
  %BJC: added following sentence to aid reader in response to Chris' question below 
  In Section~\ref{method:realizations}, we 
  explain how we use these 24 independent models in order to generate a single
  MSD prediction.
  % BJC: abridged Chris comment with respect to this paragraph here:
  % CPC: the question i have at this point as a reader in this area, given 
  % how this is setup both as predictive and clustering, is "how do they plan to
  % generate data and predict with this model if they have 24 different transition
  % mat estimates"
  % CPC: returning to comment after reading.  section 2.4 explains what you do 
  % (though i'm still fuzzy on how you estimate transition matrix probs in method 2). 
  % (BJC: Note I tried to clarify this process for method 2 below)
  % CPC_continued: in any case, you might want to avoid obvious critiques and head 
  % scratching from expert reviewers. i'd suggest moving this paragraph to end of subsection and
  % expand on above considerations after you've given more cut and dry details below and reference 2.4 stating     
  %that you say how you simulate from a model estimated one trajectory at a time via HDP-AR-HMM inference

@@

  We clustered states with like parameter sets in order to reduce the state space to
  a more easily interpretable size 
  %BJC: added following in response to Chris' below comment. I think I'm interpretting what he's saying right
  % CPC: perhaps keep hammering home not only reducing, but also reemphasize why absolute
  % value of some parameters really not of interest 
  %MRS: explain why parameters of single states don't give insight; more context needed?  Is CPC talking about
  %MRS: some some of the parmameters not being interesting in general, not just the values for some states?
  and because the parameters of a single states does not always give meaningful physical insight. 
  For each solute studied, we identified approximately 300-500 independent states, 
  each with separate VAR(1) parameters. Many of these states exhibit very similar
  dynamical behavior except their mean levels in each dimension are different.

@@

  We obtained estimates of clustered $A$ and $\Sigma$ parameters as well as a
  %MRS: did you mean to change to T to P here?
  %MRS: also, maybe ``consensus'' or ``merged'' instead of ``single''
  single $T$, describing transitions between all possible clustered states exhibited by
  the ensemble of trajectories, using the inference component of the HDP-AR-HMM
  algorithm. We applied the HDP-AR-HMM simultaneously on all 24 zeroed solute
  trajectories with state sequences fixed according to the remapped clustered 
  state assignments. 
  %MRS: be specific what ``This'' is in the sentence.
  This is only possible because the state sequence is fixed. 
  In cases where one does not know the state sequence and would like to 
  incorporate data from multiple trajectories, one can apply the beta process (BP)
  AR-HMM.~\cite{fox_sharing_2009} However the data requirements may be quite high
  for a good parameterization, which is why we elected to follow the clustering 
  approach described thus far.
  %BJC: I actually found 3 different ways do this. All giving the same results.
  % Method 1: described above - this is the most efficient way
  % Method 2: Link all 24 trajectories by adding a coffin state (basically a 
  % state that is a vector of zeros) between each traj
  % Method 3: Parameterize each trajectory independently, average together. T is
  % estimated by making a count matrix based on inputs from each trajectory

@@

  For both approaches, we drew state sequences with transition probabilities given 
  % BJC: Chris wasn't clear where T came from for method 2. I tried to make it clearer in 
  % last paragraph of previous section. Let me know if it's not
  %MRS: P vs. T agin?
  by $T$ of the HDP-AR-HMM. While in a given state, we simulated motion according to the 
  VAR(1) parameterization of that state. After each state transition, we set the 
  unconditional mean of the new state to the particle's position immediately before
  the state transition occurred.

@@

  Solute flux, as written, can be expressed in terms of the product of solute 
  diffusivity, $D_i$, within the membrane and a sorption coefficient, $K_i$, 
  that quantifies the ability of the solute to move between the bulk solution
  and and the membrane:
  \begin{equation}
    J_i = \frac{D_i K_i}{l}\Delta C
  \end{equation}
  where $l$ is the membrane thickness. For our purposes, $K_i$ and $l$ are 
  irrelevant since we do not study pore entry effects or flux across finite
  length pores. Therefore, in order to calculate selectivity, with respect
  to quantities we can study with our MD simulations, it is only necessary that
  we compare the ratio of diffusion constants:
  \begin{equation}
    S_{ij} = \frac{D_i}{D_j}
    \label{eqn:selectivity_diffusivity}
  \end{equation}
  It can be shown on time scales where motion is uncorrelated, that the diffusion
  constant is proportional to the slope of the linear region of the MSD 
  curve:~\cite{einstein_investigations_1956}
  \begin{equation}
    \overline{z^2(\tau)} = 2 n D t
    \label{eqn:msd_D}
  \end{equation}
  where $n$ is the number of spatial dimensions (1 in our case). In general, the 
  MSD of the realizations of our models enters a linear regime near a 5 ns time lag
  %MRS: ref below to figure not recognized?
  (see Figure~\ref{S-fig:MET_nclusters} of the Supporting Information for example).
  Since MSD sampling decreases with increasing time lag, longer trajectory 
  realizations help to extend the linear region.
  %BJC: need to confirm these details
  For all solutes, we generated 200 ensembles of 24 5 $\mu s$ trajectories and 
  estimated $D$ from the slope fit between 10 ns and 1000 ns. We used weighted 
  least squares to fit the slope with weights determined by the number of time 
  lags used to compute the average ($T - \tau$ in equation~\ref{eqn:tamsd}). 
  We bootstrapped error bars for the final reported selectivities. For each 
  bootstrap trial, we sampled the 200 trajectory ensembles with replacement, 
  averaging the mean MSD curves calculated from each ensemble and then calculating its 
  slope to create a distribution of $D$ values. We report the mean and standard
  error of the distribution of $D$.

@@

  %BJC6: added. Should I add a figure?
  %MRS: I think fine as supporting.
  A potential shortcoming of the HDP-AR-HMM applied to MD simulations is its 
  inability to reproduce the short time scale MSD curves. The MSDs predicted
  by the model tend to enter a linear regime much more quickly than the MD MSDs
  (see Figure~\ref{S-fig:unclustered_msd_shortlag} of the Supporting Information
  for example). This is not a major concern for predicting macroscopic properties
  since we are primarily interested in reproducing the MSDs on much longer 
  timescales.
  %BJC6: could hypothesize why using previous work as support

@@

  %MRS: I feel like this should be caveatted a bt. 
  The data suggests that the particular LLC membrane studied in this work, 
  up to the resolution of the moleceular modeling,
  is capable of selectively separating urea and acetic acid from ethylene glycol
  and methanol. Ethylene glycol and methanol have comparable diffusivities
  while urea and acetic acid have much lower diffusivities.

@@

  The parameters of the five dominant states provide clues about the associated solute
  behavior. Characteristic dynamics of the following states appear in 
  Figure~\ref{fig:clustered_traj_MET}. A more in-depth analysis of the state parameters
  is given in Section~\ref{S-section:clustered_parmeters} of the Supporting Information.
  \begin{enumerate}[label={State \theenumi :}, leftmargin=3.5\parindent]
     \item This state has intermediate dwell times with high past-hop dependence (high
     entries of $A$). The cluster's radial mean is 1.5 nm which is about halfway between
     the pore center and tail ends. Evidence from Figure~\ref{fig:clustered_traj_MET} 
     suggests that this state occurs essentially evenly across possible radial
     distances from the pore center.
     \item This state's radial mean of 2.4 nm implies that it occurs exclusively in the
     tail region based. Fluctuations are small and there is little past hop dependence.
     \item This is likely a hopping state that strongly contributes to the solute's MSD. 
     It is short lived and has large fluctuations which often results in very large hops.
     Like State 1, we see this type of behavior near the pore center and near the ends of the tails.
     \item This is a trapping state, having the longest expected dwell time among the dominant
     states, with a strong dependence on its previous fluctuation. Evidence from 
     Figure~\ref{fig:clustered_traj_MET} and a radial mean of 1.8 suggest that it tends to
     occur at intermediate distances from the pore center with perhaps a higher tendency
     to occur deeper within the tails.
     \item This is an intermediate-length trapping state that occurs 0.9 nm from the pore 
     center on average. Fluctuations are small and relatively independent of its previous 
     fluctuation.
  \end{enumerate}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{hbonds_assoc_summary.pdf}
  \caption{}\label{fig:hbonds_assoc_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{rdf_summary.pdf}
  \caption{}\label{fig:rdf_summary}
  \end{subfigure}
  %BJC: Is it worth putting a bars with the solute MSDs on this plot
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{dwell_time_summary.pdf}
  \caption{}\label{fig:dwell_time_summary}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{density_comparison.pdf}
  \caption{}\label{fig:density_comparison}
  \end{subfigure}
  \caption{(a) Solutes hydrogen bond and associate with sodium ions
  to different degrees. Sometimes, they participate in both interactions at
  once. (b) We histogrammed the radial means of the unclustered states weighted
  by the length of time spent in that state. Most solutes spend the majority of
  their time close to the pore center. Only methanol has a significant density
  far from the pore center. (c) The largest expected value (see 
  Equation~\ref{eqn:dwell_times}) of dwell times among the dominant clustered 
  states for each solute are inversely related to the solute's MSD. 
  %MRS: is the c) point actually made in the figure?  There's no MSD in that graph.  
  % BJC: Copying Chris' comments below:
  % CPC: Discussion is confusing. guessing error bars are from measured simulation 
  % dwell times in dominant state.  
  % BJC: will add error bar explanation once I figure out exactly how to modify the figure
  % CPC: what if trajectory has more than one dominant state? (seems from fig like 
  % you take one with largest dwell time from your model. 
  % BJC: All solutes have at  least 4 dominant states by my definition. maybe it would 
  % be better to sample dwell times from all of the dominant states. Just weight 
  % their contributions. Not sure if weighting should be with respect to the fraction
  % the total simulation time spent in each state, or the number of times the state
  % was visited.
  % CPC: might want to better connect MSDs of model in fig.3 to fig 9 c and 
  % expand/clarify discussion on connection of fig 9c to other panels just lots
  % of ambiguity here, but results seem promising.
  % BJC: is it worth plotting bars with the MSD after 1000 ns as part of (c)? 
  (d) The expected dwell time of solutes increases 
  in states with high local number density of surrounding heavy atoms.
  }\label{fig:summaries}
  \end{figure}

  % BJC: Chris was confused by this paragraph
  %MRS: maybe switch ``dominant'' for something else, like predominant, or consensus
  Solutes with dominant states characterized by long dwell times tend to 
  have lower MSDs. We define the dominant states exhibited by each solute
  as any state visited by at least half (12) of the solute trajectories.
  By our definition and based on our choice to study 10 clusters for each 
  solute, methanol, ethylene glycol, urea and acetic acid are characterized
  %BJC: added this supporting figure 
  %MRS: for some reason Not referencing this figure in supporting despite xr? 
  by 5, 6, 5 and 4 dominant states respectively (see Figure~\ref{fig:state_prevalence}
  of the Supporting Information). In Figure~\ref{fig:dwell_time_summary}, we
  plot the expected dwell time of each solute in the dominant state which 
  has the the highest self-transition probability. The MSDs shown in 
  Figure~\ref{fig:unclustered_msds} are inversely related to the maximum 
  expected dwell times. Both acetic acid and urea participate in many hydrogen
  bonds and sodium ion association interactions and the states visited by both 
  solutes have similar distributions of radial means, however acetic acid is 
  subject to much longer dwell times, and hence a lower MSD. This may be 
  because acetic acid is a stronger hydrogen bond donor than urea.

@@

  When clustering, our goal was to create the simplest model possible that gave
  clear mechanistic insight. This is particularly important for systems where transport mechanisms
  are not easily hypothesized. However, once one gains some mechanistic understanding,
  as we have in this work, the clustering procedure can be modified in order to 
  study specific interactions or radially dependent behavior.
